The Emperor Has No Clothes

### 2011-04-20 Winds/papers/LBG_Lya/lya_lbg.tex
profx
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5913 ac11df5e-a788-c23b-a8e1-c758c323924d

In contrast to our assertions,
other authors have modeled the \lya\ profiles of LBGs and have
concluded the redshifted emission is a `clear signature' of an outflow
\citep[e.g.][]{verhomme}.  
[page 7]

Similar to the ISM+wind model described above, these outflow models only predict a
redshift if there is a very large opacity at $v \approx 0\mkms$.  In
the \cite{verhomme} models, this is accomplished by considering a
thin expanding shell with a total \ion{H}{1} column density $\mnhi > 10^{20}
\cm{-2}$.  In this case, the damping wings of the gas imply a large
opacity at $v \approx 0 \mkms$ provided the shell has a speed less
than $\approx 500 \mkms$.  This model, then, predicts a very strong
absorption profile at $v < 0 \mkms$, which is generally not observed
in the LBG spectra.  One of the few exceptions is the galaxy cb58.]

Now consider the spatial distribution of \lya\ emission.
Figure~\ref{fig:SB} shows the predicted surface brightness profile
vs.\ projected radius for the flat-continuum intrinsic model and with
no IGM attenuation.  The signal is integrated from velocities $v =
[-100,400]\mkms$. Overplotted on the model, at an arbirtrary
normalization, is the fitted profile of S+11 to their observations,
$\mu(\rho) \propto \exp(-\rho/b)$ with $b=20$\,kpc.  In constrast to
the observations, the clump model predicts that 95\%\ of the
integrated flux is emitted at $\rho < 10$\,kpc and the model assumes
an exponential-like profile only at very large radii.  In short, a
proper radiative transfer treatment of the LBG clump model shows it
cannot reproduce the observed, extended \lya\ emission.

The explanation for the model's failure is twofold:  
(1) the covering fraction of the clumps decrease with radius to
$f_{\rm c} < 0.2$ at $\rho > 50$\,kpc; and
(2) the velocity gradient at $r>5$\,kpc is very shallow.  Therefore,
at large radii the outflow absorbs very few photons relative to the
inner regions.  In terms of the Sobolev approximation, we estimate
$\tau_{\rm S}(\rho = 60 \, {\rm kpc}) = XX \tau_{\rm S}(\rho=1\, {\rm
  kpc})$.  
Note that unlike the kinematic signatures, the inclusion of an ISM
component at small radii would not alter any of these conclusions.
[Add a last few lines]

@@

In the previous section, we demonstrated that the LBG clump model
of S+10 does not reproduce the observed \lya\ emission associated with
LBGs.   We also emphasize, however, that the radiative transfer of
\lya\ is sensitive to the intrinsic emission of the galaxy ($r \sim
10$\,kpc), the physics of the gas in the CGM ($r \sim 100$\,kpc), and
also the property of the IGM on larger scales ($r \sim 1-10$\,Mpc).
It is possible, therefore, that the observed \lya\ emission is
dominated by components other than a galactic-scale wind, indirectly
observed as blue-shifted absorption.  In this respect, the LBG clump
model cannot be ruled out by its discordance with the \lya\
observations.

@@

\begin{equation}
M_{\Delta r}(r) = 4\pi r^2 f_{\rm c}(r) \, \frac{m_{\rm c}}{\sigma}
\perd
\label{eqn:mshell}
\end{equation}
Following Sobolev formalism, we set the shell thickness with radius to
the velocity gradient of the flow

\begin{equation}
\Delta r = \Delta v_{\rm c} \ltp \frac{dv}{dr} \rtp^{-1}
\end{equation}
where $\Delta v_{\rm c}$ characterizes the velocity dispersion of the
clumps, from any combination of internal or external motions within
the shell.  In the following we adopt $\Delta v_{\rm c} = 10 \mkms$.

Lastly, we constrain the surface density of the clumps $m_{\rm
  c}/\sigma$ by demanding each clump be optically thick to the
metal-line transitions observed by S+10, e.g.\ \ion{Si}{2}~1526.  
This implies a minimum column density of Si$^+$ ions;\footnote{For this
  calculation we adopt the LBG clump model parameters from S+10 that
  apply to Si$^+$.  In this respect, the estimates may be a lower
  limit given that the Hydrogen gas is predicted to extend to larger
  radii.} 
we adopt
$\N{Si^+} > 10^{14} \cm{-2}$.  In turn, this implies a lower limit to
$m_{\rm c}/\sigma$.  For a solar abundance $n_{\rm H} = 10^{-4.5}
n_{\rm Si}$ and assuming no ionization correction, we find $N_{\rm H}
\ge 10^{18.5} \cm{-2}$ and therefore $m_{\rm c} / \sigma \ge 3.3
\sci{-2} \msun \, \rm pc^{-2}$ when one accounts for helium.

Adopting the minimum value for $\mmc/\sigma$,  we derive the mass,
mass flux, energy, and power profiles shown in
Figure~\ref{fig:mass_energy}.
[Note that the clump flux ($nvr^2$) is nearly constant (increases by a
factor of 2 somehow!) at $10^{-2}$ clumps/yr].
Regarding mass, we find the clump model has $M_{\rm TOT} \approx
10^{10} \msun$ [check] with nearly all of this matter existing at
radii $r>20$\,kpc.  Ironically, this material accounts for very little
of the down-the-barrel absorption\footnote{The only indisputable
  outflow component.} in the model: XX\%\ of the mass in the flow
corresponds to XX\%\ of the absorption.  
[Can momentum driven winds get this much energy/mass flux?]
We conclude that in their attempt to reconcile the absorption
transverse to LBGs with that along LBG sightlines that S+10 were
driven to this highly unrealistic scenario.

>>>>

In contrast to our assertions,
other authors have modeled the \lya\ profiles of LBGs and have
concluded the redshifted emission is a `clear signature' of an
outflow.
The [most extensive] modeling of \lya\ emissoin by LBGas to date has
been performed by A. Verhomme and collaborators
\citep{verhommea,verhommeb}.  Their preferred model has been a thin
expanding shell characterized by the expansion speed $v_{\rm exp}$ and
total surface density of \ion{H}{1} gas in the shell.  Although the
shell's dynamics play a role in the \lya\ emission kinematics, the
[dominant] aspect for the emergent spectrum is the \ion{H}{1} column
density.  Indeed, they only predict a dominant emission peak at $v \gg
0 \mkms$ if $\mnhi > 2\sci{20} \cm{-2}$.  For lower values, the
emission is preferentially at $v \lesssim +100 \mkms$.  The
explanation is similar to the ISM model described above;  large \nhi\
values force diffusion of \lya\ in frequency.  Furthermore, their
models predict substantial \lya\ absorption at $v < 0 \mkms$.  
Strong absorption is observed in a handful of individual spectra
\citep[e.g.][]{cb58,dessauges}, but is not evident in the stacked
spectra of \cite{shapley93} or \cite{steidel+10}.  
[Why is there sometimes weak \lya\ emission at $v < 0 \mkms$ in LAEs
at $z \sim 3$?]
[Read Shapley et al 2009 paper]
[Redshfited, asymmetric emission is suggestive of an outflow]

Now consider the spatial distribution of \lya\ emission.
Figure~\ref{fig:SB} shows the predicted surface brightness profile
vs.\ projected radius for the flat-continuum intrinsic model and with
no IGM attenuation.  The signal is integrated over velocities $v =
[-100,400]\mkms$, which corresponds to the majority of predicted line-emission.
Overplotted on the model, at an arbirtrary
normalization, is the fitted profile of S+11 to their observations,
$\mu(\rho) \propto \exp[-\rho/b]$ with $b=20$\,kpc.  
Although this model appears to be tending toward such an exponential
profile at very large impact parameter ($\rho > 50$\,kpc), the model
predict greater flux and a steeper fall-off at small radii.
In strong constrast to
the observations, the clump model predicts that 95\%\ of the
integrated flux is emitted at $\rho < 10$\,kpc.  
It is simply impossible to reconcile the LBG clump model with the
observed profile, and we further emphasize that including the IGM
attenutation and/or an intrinsically bright \lya\ emission line only
worsens the agreement.

The explanation for the model's failure is twofold:  
(1) the covering fraction of the clumps decrease with increasing radius to
$f_{\rm c} < 0.2$ at $r > 50$\,kpc; and
(2) the velocity gradient is very shallow beyond the inner few kpc. Therefore,
the outflow absorbs very few photons at large radius compared to the
inner regions.  In terms of the Sobolev approximation, we estimate
$\tau_{\rm S}(\rho = 60 \, {\rm kpc}) = 10^{-4} \tau_{\rm S}(\rho=1\, {\rm
  kpc})$.  [Figure?]
We further emphasize that the inclusion of an ISM
component at small radii would not qualitatively alter any of these conclusions.
Instead, one requires additional sources of \lya\ emission at large
radii or additional opacity in the CGM far beyond that predicgted by
the LBG clump model.  We return to these issues in
$\S$~\ref{sec:later}.

@@

In the previous section, we showed that the LBG clump model
of S+10 cannot on its own reproduce the observed \lya\ emission associated with
LBGs.   We also emphasize, however, that the radiative transfer of
\lya\ is sensitive to the intrinsic emission of the galaxy ($r \sim
10$\,kpc), the physics of the gas in the CGM ($r \sim 100$\,kpc), and
also the properties of the IGM on larger scales ($r \sim 1-10$\,Mpc).
It is possible, therefore, that the observed \lya\ emission is
dominated by components other than a galactic-scale wind.
In this respect, the LBG clump
model cannot be ruled out because of its discordance with the \lya\
observations.

@@

\begin{equation}
M_{\Delta r}(r) = 4\pi r^2 f_{\rm c}(r) \, \frac{m_{\rm c}}{\sigma}
\perd
\label{eqn:mshell}
\end{equation}
Following the Sobolev formalism, we set the shell thickness at a given
radius according to the velocity gradient of the flow,

\begin{equation}
\Delta r = \Delta v_{\rm c} \ltp \frac{dv}{dr} \rtp^{-1} \cmma
\end{equation}
where $\Delta v_{\rm c}$ characterizes the velocity dispersion of the
clumps.  This velocity dispersion may result from any combination of internal or external motions within
the shell.  In the following we adopt $\Delta v_{\rm c} = 10 \mkms$.

Lastly, we constrain the surface density of the clumps $m_{\rm
  c}/\sigma$ by demanding each clump be optically thick to the
metal-line transitions observed by S+10, e.g.\ \ion{Si}{2}~1526.  
This implies a minimum column density of Si$^+$ ions;\footnote{For this
  calculation we adopt the LBG clump model parameters from S+10 that
  apply to Si$^+$.  In this respect, the estimates may be a lower
  limit given that the Hydrogen gas is predicted to extend to larger
  radii.} 
we adopt
$\N{Si^+} > 10^{14} \cm{-2}$.  In turn, this yields a lower limit to
$m_{\rm c}/\sigma$.  For a solar abundance $n_{\rm H} = 10^{-4.5}
n_{\rm Si}$ and assuming no ionization correction, we find $N_{\rm H}
\ge 10^{18.5} \cm{-2}$ and therefore $m_{\rm c} / \sigma \ge 3.3
\sci{-2} \msun \, \rm pc^{-2}$ when one accounts for helium.

Adopting the minimum value for $\mmc/\sigma$ from above,  we derive the mass,
mass flux, energy, and power profiles shown in
Figure~\ref{fig:mass_energy}.
[Note that the clump flux ($nvr^2$) is nearly constant (increases by a
factor of 2 somehow!) at $10^{-2}$ clumps/yr].
Integrating the mass profile of the clumpy wind, we calculate 
$M_{\rm TOT} \approx 10^{10} \msun$ [check] with nearly all of this
matter laying at radii $r>20$\,kpc.  Ironically, this material accounts for very little
of the down-the-barrel absorption\footnote{The only indisputable
  outflow component.} in the model: XX\%\ of the mass in the flow
corresponds to XX\%\ of the absorption.  
[Can momentum driven winds get this much energy/mass flux?]
